{\normalsize Seminar: Formal Specification} \\[1ex] Security of Multithreaded Programs by Compilation

### 2012-02-01 paper/paper.tex
removed unecessary gabs before formulas
added \ref to figure 2
corrected footnote in motivation

The idea Barthe, Rezk, Russo and Sabelfeld phrased in \cite{Barthe07} was to close
this leak (formally called \textit{covert channel}) by annotating the byte-code
of the program\footnote{Most mobile devices use some sort of byte-code e.g. Android.} with security
labels according to the confidentiality of the information, which causes the scheduler to hide
threads that leak information from the attacker. Their formal model allows to
formulate this idea as a framework, which can be instantiated with many programming languages,
schedulers and types of byte-code.

@@

Based on this an evaluation relation $\Downarrow_{conc} \subseteq (\text{ConcState} \times \text{History}) \times
\text{GMemory})$ is defined by

\begin{align*}
s, h \Downarrow_{conc} \mu \Leftrightarrow \exists s', h':\ (s,h \leadsto_{conc}^* s',h') \land s'.act = \emptyset \land s'.gmem = \mu
\end{align*}

@@

Now we can define our goal: Noninterference. We define noninterference
in accordance to an indistinguishability relation $\sim_g$ on global memories.
Barthe et. al. state that it is not necessary -- for the purpose of the paper --
to specify the definition of this relation. But to get a feeling for this
relation, one can define it as: $\mu \sim_g \mu' \Leftrightarrow \mu|_{low} =
\mu'|_{low}$ where $\mu|_{low}$ projects out all high elements. Based on that
a program P is non-interfering if for all global memories $\mu_1, \mu_2, \mu_1'$
and $\mu_2'$ it holds that:

\begin{align*}
\mu_1 \sim_g \mu_2 \land (P,\mu_1 \Downarrow_{conc} \mu_1') \land (\text{P},\mu_2 \Downarrow_{conc} \mu_2') \Rightarrow \mu_1' \sim_g \mu_2'
\end{align*}

@@

This type system is extended by the following rules to support multithreading:

@@

The next step is to construct the \texttt{next} function. To make this task easy, we introduce
a source language which will be complied into the given assembly language. The source language
is defined as follows:

\begin{align*}
e ::= n\ |\ x\ |\ e\ op\ e && c ::= x\ :=\ e\ |\ c;c\ |\ \text{if}\ e\ \text{then}\ c\ \text{else}\ c\ |\ \text{while}\ e\ \text{do}\ c\ |\ \text{fork}(c)
\end{align*}

Now it is easy to define the control dependence region and the junction points for the source
language and deduce the ones for the assembly language from them. The first step is to label the
source language at the regions at risk with natural numbers:

\begin{align*}
c ::= [x\ :=\ e]^n\ |\ c;c\ |\ [\text{if}\ e\ \text{then}\ c\ \text{else}\ c]^n\ |\ [\text{while}\ e\ \text{do}\ c]^n\ |\ [\text{fork}(c)]^n
\end{align*}

>>>>

The idea Barthe, Rezk, Russo and Sabelfeld phrased in \cite{Barthe07} was to close
this leak (formally called \textit{covert channel}) by annotating the byte-code
of the program\footnote{Most operating systems for mobile devices use some sort of byte-code (e.g. Android) for application software.} with security
labels according to the confidentiality of the information, which causes the scheduler to hide
threads that leak information from the attacker. Their formal model allows to
formulate this idea as a framework, which can be instantiated with many programming languages,
schedulers and types of byte-code.

@@

Based on this an evaluation relation $\Downarrow_{conc} \subseteq (\text{ConcState} \times \text{History}) \times
\text{GMemory})$ is defined by
\begin{align*}
s, h \Downarrow_{conc} \mu \Leftrightarrow \exists s', h':\ (s,h \leadsto_{conc}^* s',h') \land s'.act = \emptyset \land s'.gmem = \mu
\end{align*}

@@

Now we can define our goal: Noninterference. We define noninterference
in accordance to an indistinguishability relation $\sim_g$ on global memories.
Barthe et. al. state that it is not necessary -- for the purpose of the paper --
to specify the definition of this relation. But to get a feeling for this
relation, one can define it as: $\mu \sim_g \mu' \Leftrightarrow \mu|_{low} =
\mu'|_{low}$ where $\mu|_{low}$ projects out all high elements. Based on that
a program P is non-interfering if for all global memories $\mu_1, \mu_2, \mu_1'$
and $\mu_2'$ it holds that:
\begin{align*}
\mu_1 \sim_g \mu_2 \land (P,\mu_1 \Downarrow_{conc} \mu_1') \land (\text{P},\mu_2 \Downarrow_{conc} \mu_2') \Rightarrow \mu_1' \sim_g \mu_2'
\end{align*}

@@

This type system is extended by the rules in figure~\ref{fig:multithreaded-typing-rules} to support multithreading.

@@

The next step is to construct the \texttt{next} function. To make this task easy, we introduce
a source language which will be complied into the given assembly language. The source language
is defined as follows:
\begin{align*}
e ::= n\ |\ x\ |\ e\ op\ e && c ::= x\ :=\ e\ |\ c;c\ |\ \text{if}\ e\ \text{then}\ c\ \text{else}\ c\ |\ \text{while}\ e\ \text{do}\ c\ |\ \text{fork}(c)
\end{align*}

Now it is easy to define the control dependence region and the junction points for the source
language and deduce the ones for the assembly language from them. The first step is to label the
source language at the regions at risk with natural numbers:
\begin{align*}
c ::= [x\ :=\ e]^n\ |\ c;c\ |\ [\text{if}\ e\ \text{then}\ c\ \text{else}\ c]^n\ |\ [\text{while}\ e\ \text{do}\ c]^n\ |\ [\text{fork}(c)]^n
\end{align*}
